---
title: "Kenai Peninsula Aquatic Invasive Species Survey Schedule"
format:
  gfm: 
    keep-yaml: FALSE
    toc: TRUE
    toc-depth: 2
bibliography: "data/raw_data/bibliographic_data/bibliography.bib"
link-citations: true
---

```{r setup, echo=FALSE}
sch <- read.csv("data/raw_data/planning_data/survey_schedule.csv",
 stringsAsFactors = FALSE)
sch <- sch[order(sch$year, 
 sch$waterbody_name,
 sch$target_species),]
coop <- read.csv("data/raw_data/planning_data/cooperators.csv",
 stringsAsFactors = FALSE)
coop <- coop[order(coop$name,
 coop$acronym),]
```


## Purpose

The purpose of this repository is to facilitate collaboration among multiple partners in detecting and responding to infestation of invasive aquatic species on the Kenai Peninsula, Alaska in partial fulfillment of overarching plans to address aquatic invasive species on the Kenai Peninsula [@ElodeaCommittee2021; @Dunkeretal2022].

## Cooperators

```{r cooptab, echo=FALSE}
#| label: tbl-coop
#| tbl-cap: "Kenai Peninsula aquatic invasive species survey participants."
#| tbl-cap-location: top


col.names <- c("Name",
 "Acronym",
 "URL")
knitr::kable(coop, 
  longtable = TRUE, 
  booktabs = TRUE,
  row.names = FALSE,
  col.names = col.names
)
```


## Schedule

```{r scheduletab, echo=FALSE}
#| label: tbl-schedule
#| tbl-cap: "Planned invasive species survey schedule for Kenai Peninsula waterbodies. Acronyms are the same as in @tbl-coop."
#| tbl-cap-location: top


col.names <- c("Year", 
 "Waterbody Name",
 "Latitude",
 "Longitude",
 "Target Species",
 "Surveyors")
knitr::kable(sch, 
  longtable = TRUE, 
  booktabs = TRUE,
  row.names = FALSE,
  col.names = col.names
)
```

## Planning Resources

### Prioritizations

- Alaska Invasive Species Lake Prioritization: <https://experience.arcgis.com/experience/41a6f3a3f35f4e0fae52f9c5a0c2fbd2/>

### Information about Past Surveys

- Kenai Peninsula AIS Survey and Monitoring Map:  <https://www.arcgis.com/home/item.html?id=336e4fc594ba42069995a74202cb4750> 
- Alaska Exotic Plant Information Clearinghouse database: <http://aknhp.uaa.alaska.edu/apps/akepic/>

## Methods

Waterbody names and locations from were downloaded from USGS (<https://geonames.usgs.gov/docs/stategaz/AK_Features.zip>) on March 3, 2023. We obtained planned survey schedules through personal communication and from available plans [@Massengilletal2022].

## References
