---
title: "Kenai Peninsula Aquatic Invasive Species Survey Schedule"
format:
  gfm: 
    keep-yaml: FALSE
    toc: TRUE
    toc-depth: 2
bibliography: "data/raw_data/bibliographic_data/bibliography.bib"
link-citations: true
---

```{r setup, echo=FALSE, error=FALSE, comment=FALSE, warning=FALSE, message=FALSE, results=FALSE}
# Load libraries.
library(sf)
library(zip)

# Load data.
sch <- read.csv("data/raw_data/planning_data/survey_schedule.csv",
 stringsAsFactors = FALSE)
sch <- sch[order(sch$year, 
 sch$waterbody_name,
 sch$target_species),]
coop <- read.csv("data/raw_data/planning_data/cooperators.csv",
 stringsAsFactors = FALSE)
coop <- coop[order(coop$name,
 coop$acronym),]

# Process data.
schsf <- st_as_sf(sch, coords = c("longitude", y="latitude"))
st_crs(schsf) <- "EPSG:4326"

this_year <- as.numeric(substr(Sys.time(), 1, 4))

# Export data.
st_write(schsf[schsf$year >= this_year,], 
 dsn = "data/final_data/geodata/survey_schedule.kml",
 delete_dsn = TRUE)
st_write(schsf[schsf$year >= this_year,], 
 dsn = "data/final_data/geodata/survey_schedule.shp",
 delete_dsn = TRUE)
write.csv(sch, file = "data/final_data/geodata/survey_schedule.csv", 
 row.names = FALSE)
st_write(schsf[schsf$year < this_year,], 
 dsn = "data/final_data/geodata/past_surveys.kml",
 delete_dsn = TRUE)
st_write(schsf[schsf$year < this_year,], 
 dsn = "data/final_data/geodata/past_surveys.shp",
 delete_dsn = TRUE)
if(file.exists("data/final_data/geodata/survey_schedule.zip")){
  file.remove("data/final_data/geodata/survey_schedule.zip")
}
zip::zip(zipfile = "data/final_data/geodata/survey_schedule.zip",
 files = paste0("data/final_data/geodata/", 
 dir("data/final_data/geodata")),
 mode = "cherry-pick")
```


## Purpose

The purpose of this repository is to facilitate collaboration among multiple partners for early detection of aquatic invasive species on the Kenai Peninsula, Alaska in partial fulfillment of overarching plans to address aquatic invasive species on the Kenai Peninsula [@ElodeaCommittee2021; @Dunkeretal2022].

## Cooperators

```{r cooptab, echo=FALSE}
#| label: tbl-coop
#| tbl-cap: "Kenai Peninsula aquatic invasive species survey participants."
#| tbl-cap-location: top


col.names <- c("Name",
 "Acronym",
 "URL")
knitr::kable(coop, 
  longtable = TRUE, 
  booktabs = TRUE,
  row.names = FALSE,
  col.names = col.names
)
```


## Schedule

The planned sampling schedule is available in @tbl-schedule below and it can be downloaded as a csv file in this repository at [data/raw_data/planning_data/survey_schedule.csv](data/raw_data/planning_data/survey_schedule.csv). The schedule can also be downloaded as spatial data in kml and shp formats from the [releases](https://github.com/Kenai-NWR-bio/AISschedule/releases) page.

```{r scheduletab, echo=FALSE}
#| label: tbl-schedule
#| tbl-cap: "Planned invasive species survey schedule for Kenai Peninsula waterbodies. Acronyms of surveyors are the same as in @tbl-coop."
#| tbl-cap-location: top


col.names <- c("Year", 
 "Waterbody Name",
 "Latitude",
 "Longitude",
 "Target Species",
 "Surveyors")
knitr::kable(sch[schsf$year >= this_year,], 
  longtable = TRUE, 
  booktabs = TRUE,
  row.names = FALSE,
  col.names = col.names
)
```

## Planning Resources

### Prioritizations

- Alaska Invasive Species Lake Prioritization: <https://experience.arcgis.com/experience/41a6f3a3f35f4e0fae52f9c5a0c2fbd2/>
- Finalized prioritization for Elodea surveys on the Kenai Peninsula: <https://www.arcgis.com/home/item.html?id=fe95107ac89c4d488ffa8cfdc0d65178>

### Information about Past Surveys

- Kenai Peninsula AIS Survey and Monitoring Map:  <https://www.arcgis.com/home/item.html?id=336e4fc594ba42069995a74202cb4750> 
- Alaska Exotic Plant Information Clearinghouse database: <http://aknhp.uaa.alaska.edu/apps/akepic/>

Some data on past surveys have been gathered into @tbl-past.

```{r scheduletab, echo=FALSE}
#| label: tbl-past
#| tbl-cap: "Past aquatic invasive species surveys in Peninsula waterbodies. Acronyms of surveyors are the same as in @tbl-coop."
#| tbl-cap-location: top


col.names <- c("Year", 
 "Waterbody Name",
 "Latitude",
 "Longitude",
 "Target Species",
 "Surveyors")
knitr::kable(sch[schsf$year < this_year,], 
  longtable = TRUE, 
  booktabs = TRUE,
  row.names = FALSE,
  col.names = col.names
)
```


## Methods

We obtained planned survey schedules through personal communication and from available plans [@Massengilletal2022]. When not provided, we obtained names and locations of waterbodies from USGS GNIS (<https://geonames.usgs.gov/docs/stategaz/AK_Features.zip>) or Google Earth (<https://earth.google.com/>).

Data were processed using R, version 4.2.2 [@RCoreTeam2022] and the R packages readxl, version 1.4.1 [@WickhamBryan2022]; sf, version 1.0-9 [@Pebesma2018]; and zip, version 2.2.2 [@Csardietal2022].

## References
